A Hierarchical Bayesian Framework for Stellar Obliquity Distribution Inference

### 2023-02-14 src/tex/ms.tex
[showyourwork] overleaf sync

Stellar obliquity $\psi$, the angle between a planet's orbital axis and its host star's spin axis, traces the formation and evolution of a planetary system. In transiting-exoplanet observations, only the sky-projected stellar obliquity $\lambda$ can be measured. To find the true stellar obliquity $\psi$, information about stellar inclination $i_\star$ is also required.
Here we show that, while the stellar inclination is important to infer the stellar obliquity of an individual system, they are not necessarily required for the population stellar obliquity inference. 
Constraints on the stellar obliquity distribution are predominantly contributed by the sky-projected stellar obliquities.
We introduce a flexible, hierarchical Bayesian framework that could infer the stellar obliquity distribution purely from sky-projected stellar obliquities. Stellar inclination measurements are optional in the model, and if not provided, are assumed to be isotropically distributed.
Applying the framework to all exoplanetary systems with sky-projected stellar obliquity measurements, we find that the stellar obliquity distribution is unimodal and peaked at zero degrees. Misaligned systems have nearly isotropic stellar obliquities, which may have important implications for close-in planets' formation and evolution.

@@

Stellar obliquity $\psi$ describes the angle between a planet's orbital axis $\bf{n}_{\rm orb}$ and its host star's spin axis $\bf{n}_{\star}$. 
The angle is an important tracer of a planetary system's formation environment and dynamical evolution. The evolution of stellar obliquity can be roughly summarized into three stages. First, the formation and evolution of a protoplanetary disk around its host star determine the primordial stellar obliquity \citep[e.g.,][]{Bate10, Lai11, Batygin12}. Second, post-formation dynamical evolution in the planetary system, such as planet-planet scattering \citep[e.g.,][]{Rasio96, Chatterjee08, Nagasawa08, Beague12}, von Zeipel-Kozai-Lidov mechanisms \citep[e.g.,][]{Wu03, Naoz16}, and secular chaos \citep{Wu11}, excites the mutual inclinations between planetary or stellar companions, further sculpting the stellar obliquity. Lastly, tidal realignment of the host star's spin axis to the planet's orbital axis could reduce the stellar obliquity, if the tidal dissipation in the star is efficient \citep[e.g.,][]{Winn10, Albrecht12}. Furthermore, massive stars with convective cores could generate internal gravity waves and dissipate angular momentum to their radiative zones, modifying the stellar obliquity \citep{Rogers12, Rogers13}.

It is yet unclear if all of these physical and dynamic processes apply to exoplanetary systems. These proposed mechanisms have different predictions on stellar obliquity distributions \citep[see][and references therein]{Albrecht22}. For example, in the secular chaos mechanism, the stellar obliquity could never get beyond $90\degr$ \citep{Teyssandier19}. In the stellar Kozai mechanism, the stellar obliquity distribution should be bimodal and concentrated at $40\degr$ and $140\degr$ \citep[e.g.,][]{Fabrycky07, Anderson16, Vick19}. In the multiple-planet scattering mechanism, the majority of systems should be aligned, and a small fraction of systems have a wide range of stellar obliquities \citep[e.g.,][]{Beague12}. Motivated by these predictions, we characterize the stellar obliquity distribution of exoplanetary systems from a Bayesian approach to find the dominant mechanisms that sculpt close-in planetary systems.

In exoplanetary observation, only the sky-projected stellar obliquity $\lambda$, the angle between the projections of $\bf{n}_{\rm orb}$ and $\bf{n}_{\star}$ onto the plane of the sky, can be measured, mainly via the Rossiter-McLaughlin effect \citep{Rossiter24, McLaughlin24}. The stellar obliquity $\psi$ can be calculated, if both the sky-projected stellar obliquity $\lambda$ and the stellar inclination $i_\star$ are known. The relation follows \citep[e.g.,][]{Fabrycky09}:
\begin{equation}\label{eqn:psi}
    \cos{\psi} = \sin{i_\star}\sin{i_{\rm orb}}\cos{\lambda} + \cos{i_\star}\cos{i_{\rm orb}},
\end{equation}
where $i_{\rm orb}$ is the inclination angle between the vector $\bf{n}_{\rm orb}$ and the observer's line of sight, and $i_\star$ is the inclination angle between $\bf{n}_{\star}$ and the observer's line of sight.
For transiting exoplanet systems where $i_{\rm orb} \approx 90\degr$, $\cos{\psi} \approx \sin{i_\star}\cos{\lambda}$.

Stellar inclinations can often be constrained via, e.g., photometric and spectroscopic rotational modulation introduced by starspots \citep[e.g.,][]{Masuda20, Albrecht21}, gravity darkening \citep[e.g.,][]{Barnes09, Barnes11}, and asteroseismology \citep[e.g.][]{Chaplin13}.
For systems without $i_\star$ measurements, it is still possible to infer their stellar obliquities from the sky-projected obliquities, assuming isotropic stellar inclinations; however, the inferred $\psi$ will have greater uncertainty than the one inferred with $i_\star$ measurement \citep{Fabrycky09}.

We do not yet understand the role of stellar inclination in population stellar obliquity inference.
In this work, we study the relationship between the distributions of stellar obliquity, sky-projected stellar obliquity, and stellar inclination, and introduce a statistical approach to infer the stellar obliquity distribution.
In Section~\ref{sec:hbm}, we introduce a flexible, hierarchical Bayesian framework that infers the stellar obliquity distribution of a sample.
In Section~\ref{sec:applications}, we apply the framework to simulated data and real observations and show that the stellar obliquity distribution inferred purely from sky-projected stellar obliquities will not compromise our interpretation.
In Section~\ref{sec:jacobian}, we find the mathematical relation between the $\psi$, $\lambda$, and $i_\star$ distributions and explain why the stellar obliquity distribution inference is predominantly determined by sky-projected stellar obliquities.

@@

To find the stellar obliquity distribution of exoplanetary systems, we build a hierarchical Bayesian framework that takes the sky-projected stellar obliquity $\lambda$ and orbital inclination $i_{\rm orb}$ as observed data. Measurements on the stellar inclination $i_\star$ are optional. If the $i_\star$ information is not provided, we simply assume that they follow an isotropic distribution. Otherwise, the observed $i_\star$ could be taken from previous measurements, or if the rotational modulation method is applied, the $i_\star$ could be inferred from the stellar rotation period $P_{\rm rot}$, stellar radius $R_\star$, and sky-projected rotational broadening velocity $v\sin{i}_\star$.

\begin{figure}[ht!]
    \script{graph.py}
    \includegraphics[width=\linewidth]{figures/graph.pdf}
    \caption{Hierarchical Bayesian framework to infer the stellar obliquity distribution of exoplanetary systems. The stellar obliquity distribution is described by hyperparameters $\vb*{\beta}$ and constrained by $n_{\rm th}$ system's stellar obliquity $\psi_n$. Each $\psi_n$ is calculated by the system's sky-projected stellar obliquity $\lambda_n$, orbital inclination $i_{{\rm orb}, n}$, and stellar inclination $i_{\star, n}$, if available. $\gamma_{\star, n}$ contains properties of the star other than inclination, such as stellar radius and rotation period for $i_{\star, n}$ inference. Obs$_{\star,n}$ includes all the observed properties of the star with uncertainties, $\hat{\lambda}_n$ has the measured sky-projected stellar obliquity and its uncertainty, and $\hat{i}_{{\rm orb}, n}$ has measured orbital inclination and its uncertainty.}
    \label{fig:graph}
\end{figure}

Figure~\ref{fig:graph} shows the probabilistic graphical model for the hierarchical Bayesian framework. We aim to constrain a set of hyperparameters $\bm{\beta}$ that describes the stellar obliquity $\psi$ distribution. The parameter set $\bm{\beta}$ is constrained by $N$ individual systems in which $\psi_n$ is a random variable. $\psi_n$ is computed from the sky-projected stellar obliquity $\lambda_n$, the orbital inclination $i_{{\rm orb},n}$, and the stellar inclination $i_{\star, n}$. 
The parameter $\gamma_{\star, n}$ is a parameter set that contains all the stellar properties other than $i_{\star,n}$, such as the stellar rotation period $P_{{\rm rot},n}$, radius $R_{\star,n}$, and projected rotational velocity $v\sin{i_{\star,n}}$, if they are available.

The sky-projected stellar obliquity $\lambda_n$ is constrained by observed $\hat{\lambda}_n$ via the Rossiter-McLaughlin effect or gravity darkening.
The orbital inclination $i_{{\rm orb},n}$ is constrained by the transit light curves of exoplanets.
Obs$_{\star,n}$ includes observed data that helps to constrain $i_{\star,n}$ and $\gamma_{\star, n}$.
If the rotational modulation method is used to constrain the stellar inclination, the $i_{\star,n}$ and $\gamma_{\star,n}$ are constrained by rotational modulation in photometric or spectroscopic time series and observed sky-projected stellar rotational line broadening $\hat{v}\sin{i_{\star,n}}$ \citep[e.g.,][]{Masuda20}, where we have $v\sin{i_\star} = 2 \pi R_\star / P_{\rm rot}$.
If the gravity-darkening method is used, the $i_{\star,n}$ could be constrained by the anomaly in transit light curves \citep{Barnes09}. If asteroseismology is used, the $i_{\star,n}$ could be constrained from the mode splitting in the Fourier transform of photometric time series \citep{Chaplin13}.

For the stellar obliquity distribution, we model the $\cos{\psi}$ distribution instead of $\psi$ distribution to understand if the stellar obliquity is isotropically distributed. If stellar obliquity distribution follows an isotropic distribution, the $\cos{\psi}$ distribution will be uniformly distributed between -1 to 1. We adopt a two-component mixture Beta distribution with hyperparameters $\bm{\beta} = \{\bm{w},\bm{a},\bm{b}\}$, where each hyperparameter has a dimension of 2. One component is designed for a large population of well-aligned systems, and the other is designed for misaligned systems. The Beta distribution is a flexible distribution that can describe the spike of aligned systems with stellar obliquities close to $0\degr$ or the clustered or broadly distributed misaligned systems.
The probability density function of $\cos{\psi}$ follows 
\begin{align}
    \cos{\psi} &\sim 2\times\Bigl( w_0 {\rm Beta}(a_0, b_0) + w_1 {\rm Beta}(a_1, b_1)\Bigr)-1.
\end{align}
Since the Beta distribution is defined on the interval $[0, 1]$, we extend its support to $[-1,1]$ using a linear transformation (i.e., $2\times {\rm Beta} - 1$). For the ${\rm Beta}(a_0,b_0)$ component, we have it to describe the population of aligned systems. For the $a_0/b_0 \gg 1$, the distribution will have a spike at $\cos{\psi} = 1$. For the ${\rm Beta}(a_1,b_1)$ component, we have it to describe the misaligned systems. We design the order of the two components to avoid label switching in the mixture model.
With the design, the hyperparameter and parameter priors are the following:
\begin{align}
    w_{0,1} &\sim {\rm Dirichlet}(1, 1) \nonumber\\
    a_0 &\sim \mathcal{U}(0, 50) \nonumber\\
    b_0 &\sim \mathcal{U}(0, 1) \nonumber\\
    a_1 &\sim \mathcal{U}(0, 10) \nonumber\\
    b_1 &\sim \mathcal{U}(0, 10) \nonumber\\
    \cos{i}_{\star,n} &\sim \mathcal{U}(0, 1) \nonumber\\
    \cos{i}_{{\rm orb},n} &\sim \mathcal{U}(-1, 1) \nonumber\\
    \lambda_{\star,n} &\sim \mathcal{U}(0, \pi).
\end{align}
The likelihood functions follow:
\begin{align}
    \mathcal{L}(\lambda) &\sim \prod_{i=1}^N\mathcal{N}(\hat{\lambda}_n, \sigma_{\hat{\lambda}_n}) \nonumber\\
    \mathcal{L}(i_{\rm orb}) &\sim \prod_{i=1}^N\mathcal{N}(\hat{i}_{{\rm orb},n}, \sigma_{\hat{i}_{{\rm orb},n}}) \nonumber\\
    \mathcal{L}(i_\star) &\sim \prod_{i=1}^N\mathcal{N}({\rm Obs}_{\star,n}, \sigma_{{\rm Obs}_{\star,n}}).
\end{align}
If $\gamma_{\star, n}$ is available, we construct Normal distributions with means and standard deviations from their observed data.
Here we assume uniform hyperpriors for the Beta distribution. For population inference on a small number of systems (i.e., $N \lesssim 50$), the choice of hyperpriors could impact the inferred distribution \citep{Nagpal22}. When applying the framework to a small sample size, it is important to test the robustness of inferred distributions on different hyperpriors.

The probabilistic model is built with the $\mathtt{PyMC}$ \citep[$\mathtt{v4.1.7}$;][]{pymc} package and the posteriors are sampled using the No-U-Turn Sampler \citep[NUTS;][]{Hoffman11}, a gradient-based sampling algorithm of the Markov chain Monte Carlo (MCMC). All figures and simulations in this paper are fully reproducible and built with the \showyourwork package. The open source code can be found on \href{https://github.com/jiayindong/obliquity}{GitHub\,\faGithub\,(https://github.com/jiayindong/obliquity)}.

\section{Applications to Simulated and Observational Data}\label{sec:applications}

@@

In Figure~\ref{fig:simulation}, we present the inferred stellar obliquity distributions. Since the simulated stellar obliquity distributions only have a single component, we model the data with a single Beta distribution. Each row of Figure~\ref{fig:simulation} corresponds to an injected stellar obliquity distribution. The orange curve and contours are the median and 1-$\sigma$ and 2-$\sigma$ uncertainties of the inferred $\cos{\psi}$ distribution with stellar inclination information, and the blue curve and contours are the ones without stellar inclination information. We simply assume an isotropic distribution of $i_\star$ for these models.
Surprisingly, in all four distributions, the inferred $\cos{\psi}$ distributions with or without $i_\star$ correctly recover the injected distributions shown as grey dashed lines as shown in Figure~\ref{fig:simulation}. The inferred modes and widths of the stellar obliquity distributions with or without $i_\star$ likelihood are consistent, despite that the inferred distributions without $i_\star$ measurements could have wider uncertainties shown in shallow color contours.
Since the injected distributions are Normal distributions in rows 2--4, it is expected that the inferred distributions, which as Beta distributions, could not exactly describe the injected distributions.

We also examine the importance of orbital inclination $i_{\rm orb}$ to the stellar obliquity distribution inference. Since all the systems we study are transiting-planet systems, we consider an isotropic orbital inclination distribution between $80\degr$ and $90\degr$. This broad inclination range corresponds to an impact parameter range from 0 to 1 with a planet-star separation $a/R_\star$ of 6.
We compare the inferred stellar obliquity distribution by approximating $i_{\rm orb}$ to $90\degr$ with the inferred distribution using the true $i_{\rm orb}$, and find the difference between the two distributions is minimal. For transiting-planet systems, approximating orbital inclinations to $90\degr$ will not compromise the stellar obliquity distribution inference.

From the simulated data, we confirm that while stellar inclinations help to constrain the stellar obliquity of individual systems, they are not necessarily required for the population stellar obliquity inference. The constraint on stellar obliquity distribution is predominantly based on sky-projected stellar obliquities. Just from the sky-projected stellar obliquity, we could correctly infer the peak and width of the underlying stellar obliquity distribution. Later in Section~\ref{sec:jacobian}, we derive the Jacobian transformations between $\psi$, $\lambda$, and $i_\star$ to explain why there is no significant information loss when not having $i_\star$ measurements.

\begin{figure}[ht!]
    \script{sim_plot.py}
    \includegraphics[width=\linewidth]{figures/simulation.pdf}
    \caption{Inferred stellar obliquity distributions from sky-projected stellar obliquities with or without stellar inclination measurements. Each row presents a set of simulated data where the underlying $\cos{\psi}$ distribution is known and shown as grey dashed curves. The inferred $\cos{\psi}$ distributions with or without stellar inclination measurements are shown in orange and blue curves, respectively. The shallow contours are 1-$\sigma$ and 2-$\sigma$ uncertainties of the inferred distributions.}
    \label{fig:simulation}
\end{figure}

@@

We next apply the hierarchical Bayesian framework to all exoplanetary systems with sky-projected stellar obliquity $\lambda$ measurements. We use the sample of 161 systems summarized in \cite{Albrecht22} Table A1, which are mainly Hot Jupiter systems. The inferred $\cos{\psi}$ distribution is shown in Figure~\ref{fig:psi_dist}. The $\cos{\psi}$ distribution peaks at 1 but is nearly flat between $-0.75$ and $0.75$ with no significant clustering. The distribution corresponds to a pile-up of planets with stellar obliquity less than $40\degr$ and an isotropic distribution of planets with stellar obliquity from $40\degr$ to $140\degr$. 
The aligned-system population dominates the distribution with a fraction of $w_0 = 0.67 \pm 0.09$. The hyperposteriors of the Beta distribution are $a_0 = 31.0\pm12.8$ and $b_0 = 0.40\pm0.12$.
The misaligned-system population has a fraction of $w_1 = 0.33 \pm 0.09$ and hyperposteriors of $a_1 = 1.56\pm0.93$ and $b_1 = 1.64\pm1.58$.

\begin{figure}[ht!]
    \script{psi_plot.py}
    \begin{centering}
        \includegraphics{figures/psi_dist.pdf}
        \caption{Stellar obliquity distribution of all exoplanetary systems with sky-projected stellar obliquity measurements. The inference is purely from observed sky-projected stellar obliquities and assumes the stellar inclination is isotropic.}
        \label{fig:psi_dist}
    \end{centering}
\end{figure}

The inferred stellar obliquity distribution of the full sample of systems is at odds with the previous analysis of the subsample of systems that also have $i_\star$ measurements. The subsample identified a preponderance of perpendicular planets that disfavored an isotropic stellar obliquity distribution \citep{Albrecht21}. We speculate two possible explanations for this observation discrepancy, which are open to further investigation.
First, the sample of systems with $i_\star$ measurements is small, and only about 20 systems are misaligned systems used to constrain the misaligned systems' stellar obliquity distribution. Second, the sample could have a selection bias introduced by the $i_\star$ measurement requirement.

>>>>

Stellar obliquity $\psi$, the angle between a planet's orbital axis and its host star's spin axis, traces the formation and evolution of a planetary system. In transiting-exoplanet observations, only the sky-projected stellar obliquity $\lambda$ can be measured. To determine the true stellar obliquity $\psi$, information on the stellar inclination $i_\star$ is also needed.
Here we find that while the stellar inclination is crucial for understanding the obliquity of individual systems, it is not a requirement for determining the population-level stellar obliquity. 
The constraints on the stellar obliquity distribution are mainly driven by the sky-projected stellar obliquities.
We introduce a flexible, hierarchical Bayesian framework that can infer the stellar obliquity distribution solely from sky-projected stellar obliquities. Stellar inclination measurements are optional input in the model, and if not available, they are assumed to be isotropically distributed.
When applying the framework to all exoplanetary systems with sky-projected stellar obliquity measurements, we find that the stellar obliquity distribution is unimodal and peaked at zero degrees. Misaligned systems have nearly isotropic stellar obliquities, which could have important implications for the formation and evolution of close-in planets.

@@

Stellar obliquity $\psi$ describes the angle between a planet's orbital axis $\bf{n}_{\rm orb}$ and its host star's spin axis $\bf{n}_{\star}$. 
This angle is an important tracer of a planetary system's formation environment and dynamical evolution. The evolution of stellar obliquity can be roughly broken down into three stages. First, the formation and evolution of a protoplanetary disk determine the primordial stellar obliquity \citep[e.g.,][]{Bate10, Lai11, Batygin12}. Second, post-formation dynamical evolution in the planetary system, such as planet-planet scattering \citep[e.g.,][]{Rasio96, Chatterjee08, Nagasawa08, Beague12}, von Zeipel-Kozai-Lidov mechanisms \citep[e.g.,][]{Wu03, Naoz16}, and secular chaos \citep{Wu11}, can excite the mutual inclinations between planetary or stellar companions and alter the stellar obliquity. Lastly, the tidal force can reduce the stellar obliquity by realigning the host star's spin axis with the planet's orbital axis, if the tidal dissipation in the star is efficient \citep[e.g.,][]{Winn10, Albrecht12}. Additionally, massive stars with convective cores could generate internal gravity waves and dissipate angular momentum to their radiative zones, potentially affecting the stellar obliquity \citep{Rogers12, Rogers13}.

It is yet unclear if all of these physical and dynamic processes apply to exoplanetary systems. These proposed mechanisms have different predictions on stellar obliquity distributions \citep[see][and references therein]{Albrecht22}. For example, the secular chaos mechanism suggests that the stellar obliquity would never exceed $90\degr$ \citep[e.g.,][]{Teyssandier19}. The stellar Kozai mechanism predicts that the stellar obliquity distribution should be bimodal and concentrated at $40\degr$ and $140\degr$ \citep[e.g.,][]{Fabrycky07, Anderson16, Vick19}. The multiple-planet scattering mechanism proposes that the majority of systems would be aligned, while a small fraction of systems would exhibit a diverse range of stellar obliquities \citep[e.g.,][]{Beague12}. 
With these predictions in mind, we aim to determine the dominant mechanisms responsible for shaping close-in planetary systems by characterizing the stellar obliquity distribution of exoplanetary systems through a Bayesian approach.

In exoplanetary observation, only the sky-projected stellar obliquity $\lambda$, the angle between the projections of $\bf{n}_{\rm orb}$ and $\bf{n}_{\star}$ onto the plane of the sky, can be measured. This measurement is primarily obtained via the Rossiter-McLaughlin effect \citep{Rossiter24, McLaughlin24}. The stellar obliquity $\psi$ can be determined, if both the sky-projected stellar obliquity $\lambda$ and the stellar inclination $i_\star$ are known. The relationship between $\psi$ and $\{\lambda, i_\star\}$ is given by \citep[e.g.,][]{Fabrycky09}:
\begin{equation}\label{eqn:psi}
    \cos{\psi} = \sin{i_\star}\sin{i_{\rm orb}}\cos{\lambda} + \cos{i_\star}\cos{i_{\rm orb}},
\end{equation}
where $i_{\rm orb}$ is the inclination angle between the vector $\bf{n}_{\rm orb}$ and the observer's line of sight, and $i_\star$ is the inclination angle between $\bf{n}_{\star}$ and the observer's line of sight.
In transiting-exoplanet systems where $i_{\rm orb} \approx 90\degr$, $\cos{\psi} \approx \sin{i_\star}\cos{\lambda}$.

Stellar inclinations can often be constrained via, e.g., photometric and spectroscopic rotational modulation introduced by starspots for cool stars \citep[e.g.,][]{Masuda20, Albrecht21}, gravity darkening for fast-rotating stars \citep[e.g.,][]{Barnes09, Barnes11}, and asteroseismology for bright stars \citep[e.g.][]{Chaplin13}.
For systems without $i_\star$ measurements, it is still possible to infer their stellar obliquities from the sky-projected obliquities, assuming isotropic stellar inclinations; however, the inferred $\psi$ will have greater uncertainty than the one inferred with $i_\star$ measurement \citep{Fabrycky09}.

The relationship between the distributions of stellar obliquity, sky-projected stellar obliquity, and stellar inclination is still not fully understood. In this study, we aim to gain a deeper understanding of this relationship and develop a statistical approach to infer the stellar obliquity distribution.
In Section~\ref{sec:hbm}, we introduce a flexible, hierarchical Bayesian framework that allows us to infer the stellar obliquity distribution of a sample.
In Section~\ref{sec:applications}, we apply the framework to simulated data and real observations and show that the inferred stellar obliquity distribution from sky-projected stellar obliquities is robust even if the $i_\star$ information is not provided.
In Section~\ref{sec:jacobian}, we further investigate the relationship between the distributions of $\psi$, $\lambda$, and $i_\star$, and explain why sky-projected stellar obliquities play a predominant role in determining the inferred distribution of stellar obliquities.

@@

To find the stellar obliquity distribution of exoplanetary systems, we develop a hierarchical Bayesian framework that takes the observed sky-projected stellar obliquity $\lambda$ and orbital inclination $i_{\rm orb}$ as input data. The measurement of the stellar inclination $i_\star$ is optional. 
In the absence of $i_\star$ information, we assume that it follows an isotropic distribution.
Otherwise, the observed $i_\star$ can be taken from previous measurements, or if the rotational modulation method is employed, the $i_\star$ value can be inferred from the stellar rotation period $P_{\rm rot}$, stellar radius $R_\star$, and sky-projected rotational broadening velocity $v\sin{i}_\star$.

\begin{figure}[ht!]
    \script{graph.py}
    \includegraphics[width=\linewidth]{figures/graph.pdf}
    \caption{Hierarchical Bayesian framework to infer the stellar obliquity distribution of exoplanetary systems. The distribution is described by hyperparameters $\vb*{\beta}$ and is constrained by the sky-projected stellar obliquity $\lambda_n$, orbital inclination $i_{{\rm orb}, n}$, and, if available, the stellar inclination $i_{\star, n}$ of each $n_{\rm th}$ individual system. The system's stellar obliquity $\psi_n$ is calculated based on these parameters. 
    Properties of the star other than its inclination, such as its radius and rotation period, are represented by $\gamma_{\star, n}$. 
    Obs$_{\star,n}$ contains all observed properties of the star with uncertainties, $\hat{\lambda}_n$ has the measured sky-projected stellar obliquity and its uncertainty, and $\hat{i}_{{\rm orb}, n}$ has measured orbital inclination and its uncertainty.}
    \label{fig:graph}
\end{figure}

Figure~\ref{fig:graph} illustrates the probabilistic graphical model for our hierarchical Bayesian framework. We aim to constrain a set of hyperparameters $\bm{\beta}$ that describes the stellar obliquity distribution. The parameter set $\bm{\beta}$ is constrained by $N$ individual systems, where each $\psi_n$ is computed from the system's sky-projected stellar obliquity $\lambda_n$, the orbital inclination $i_{{\rm orb},n}$, and if available, the stellar inclination $i_{\star, n}$. 
The parameter $\gamma_{\star, n}$ contains all stellar properties other than $i_{\star,n}$, such as the stellar rotation period $P_{{\rm rot},n}$, radius $R_{\star,n}$, and projected rotational velocity $v\sin{i_{\star,n}}$, if they are known.

The sky-projected stellar obliquity $\lambda_n$ is constrained by observed $\hat{\lambda}_n$ via the Rossiter-McLaughlin effect or gravity darkening.
The orbital inclination $i_{{\rm orb},n}$ of exoplanets is inferred from their transit light curves.
Obs$_{\star,n}$ includes observed data that helps to constrain $i_{\star,n}$ and $\gamma_{\star, n}$.
If the rotational modulation method is used to constrain the stellar inclination, the $i_{\star,n}$ and $\gamma_{\star,n}$ are constrained by rotational modulation in photometric or spectroscopic time series and observed sky-projected stellar rotational line broadening $\hat{v}\sin{i_{\star,n}}$ \citep[e.g.,][]{Masuda20}, where we have $v\sin{i_\star} = 2 \pi R_\star / P_{\rm rot}$.
If the gravity-darkening method is used, the $i_{\star,n}$ could be constrained by the anomaly in transit light curves \citep{Barnes09}. If asteroseismology is used, the $i_{\star,n}$ could be constrained from the mode splitting in the Fourier transform of photometric time series \citep{Chaplin13}.

For the stellar obliquity distribution, we model the $\cos{\psi}$ distribution instead of $\psi$ distribution to understand if the stellar obliquity is isotropically distributed. If stellar obliquity distribution follows an isotropic distribution, the $\cos{\psi}$ distribution will be uniformly distributed between -1 to 1. We adopt a two-component mixture Beta distribution with hyperparameters $\bm{\beta} = \{\bm{w},\bm{a},\bm{b}\}$, where each hyperparameter has a dimension of 2. One component is designed for a large population of well-aligned systems, and the other is designed for misaligned systems. The Beta distribution is a flexible distribution that can describe the spike of aligned systems with stellar obliquities close to $0\degr$ or the clustered or broadly distributed misaligned systems.
The probability density function of $\cos{\psi}$ follows 
\begin{align}
    \cos{\psi} &\sim 2\times\Bigl( w_0 {\rm Beta}(a_0, b_0) + w_1 {\rm Beta}(a_1, b_1)\Bigr)-1.
\end{align}
Since the Beta distribution is defined on the interval $[0, 1]$, we extend its support to $[-1,1]$ using a linear transformation (i.e., $2\times {\rm Beta} - 1$). For the ${\rm Beta}(a_0,b_0)$ component, we have it to describe the population of aligned systems. For the $a_0/b_0 \gg 1$, the distribution will have a spike at $\cos{\psi} = 1$. For the ${\rm Beta}(a_1,b_1)$ component, we have it to describe the misaligned systems. We design the order of the two components to prevent label switching in the mixture model.
With the design, the hyperparameter and parameter priors are the following:
\begin{align}
    w_{0,1} &\sim {\rm Dirichlet}(1, 1) \nonumber\\
    a_0 &\sim \mathcal{U}(0, 50) \nonumber\\
    b_0 &\sim \mathcal{U}(0, 1) \nonumber\\
    a_1 &\sim \mathcal{U}(0, 10) \nonumber\\
    b_1 &\sim \mathcal{U}(0, 10) \nonumber\\
    \cos{i}_{\star,n} &\sim \mathcal{U}(0, 1) \nonumber\\
    \cos{i}_{{\rm orb},n} &\sim \mathcal{U}(-1, 1) \nonumber\\
    \lambda_{\star,n} &\sim \mathcal{U}(0, \pi).
\end{align}
The likelihood functions follow:
\begin{align}
    \mathcal{L}(\lambda) &\sim \prod_{i=1}^N\mathcal{N}(\hat{\lambda}_n, \sigma_{\hat{\lambda}_n}) \nonumber\\
    \mathcal{L}(i_{\rm orb}) &\sim \prod_{i=1}^N\mathcal{N}(\hat{i}_{{\rm orb},n}, \sigma_{\hat{i}_{{\rm orb},n}}) \nonumber\\
    \mathcal{L}(i_\star) &\sim \prod_{i=1}^N\mathcal{N}({\rm Obs}_{\star,n}, \sigma_{{\rm Obs}_{\star,n}}).
\end{align}
If $\gamma_{\star, n}$ is available, we construct Normal distributions with means and standard deviations from measurements.
Here we use uniform hyperpriors for the Beta distribution. However, it is important to note that when applying the framework to a small sample size with $N \lesssim 50$, the choice of hyperpriors could impact the inferred distribution \citep{Nagpal22}. To ensure the robustness of the inferred distributions in such cases, it is crucial to test their sensitivity to different hyperpriors.

The probabilistic model is constructed using the $\mathtt{PyMC}$ package version $\mathtt{v4.1.7}$ \citep{pymc}, and the posteriors are sampled with the No-U-Turn Sampler \citep[NUTS;][]{Hoffman11}, which is a gradient-based Markov chain Monte Carlo (MCMC) sampling algorithm. This paper's figures and simulations are completely reproducible and were created using the \showyourwork package. The open-source code is available on \href{https://github.com/jiayindong/obliquity}{GitHub\,\faGithub\,(https://github.com/jiayindong/obliquity)}.

\section{Applications to Simulated and Observed Data}\label{sec:applications}

@@

In Figure~\ref{fig:simulation}, we present the inferred stellar obliquity distributions. Since the simulated stellar obliquity distributions only have a single component, we model the data with a single Beta distribution. Each row of Figure~\ref{fig:simulation} corresponds to an injected stellar obliquity distribution. The orange curve and contours are the median and 1-$\sigma$ and 2-$\sigma$ uncertainties of the inferred $\cos{\psi}$ distribution with stellar inclination information, and the blue curve and contours are the ones without stellar inclination information. We simply assume an isotropic distribution of $i_\star$ for these models.
Surprisingly, as shown in Figure~\ref{fig:simulation}, the inferred distributions of $\cos{\psi}$, whether or not they include the information of $i_\star$, correctly recover the injected distributions (represented by grey dashed lines) in all four cases. Despite the fact that the inferred distributions without $i_\star$ measurements may have wider uncertainties, as indicated by shallow color contours, the modes and widths of the inferred stellar obliquity distributions are consistent with or without $i_\star$ likelihood.
Since the injected distributions in rows 2-4 are Normal distributions, it should not be surprising that the inferred distributions, which are Beta distributions, may not exactly match the injected distributions.

We also examine the role of orbital inclination $i_{\rm orb}$ in the stellar obliquity distribution inference. Since our study focuses on transiting-exoplanet systems, we consider an isotropic orbital inclination distribution between $80\degr$ and $90\degr$. This broad range of inclinations corresponds to an impact parameter range from 0 to 1 with a planet-star separation $a/R_\star$ of 6.
We compare the stellar obliquity distributions obtained by approximating $i_{\rm orb}$ to $90\degr$ with the distributions obtained using the actual $i_{\rm orb}$. We find the difference between the two distributions is negligible. This suggests that for transiting-exoplanet systems, approximating orbital inclinations as $90\degr$ will not compromise the stellar obliquity distribution inference.

We demonstrate through simulations that the inferred stellar obliquity distribution is robust even if the $i_\star$ information is not provided, and the $i_\star$ measurement only improves the constraint on the stellar obliquity distribution. Later in Section~\ref{sec:jacobian}, we derive the Jacobian transformations between $\psi$, $\lambda$, and $i_\star$ to understand why sky-projected stellar obliquities play a predominant role in determining the inferred distribution of stellar obliquities.

\begin{figure}[ht!]
    \script{sim_plot.py}
    \includegraphics[width=\linewidth]{figures/simulation.pdf}
    \caption{Inferred stellar obliquity distributions from sky-projected stellar obliquities with and without information on the stellar inclination, depicted in blue and orange curves, respectively. Each row presents a set of simulated data with the true distribution of $\cos{\psi}$ indicated by grey dashed curves. The shallow contours represent the 1-$\sigma$ and 2-$\sigma$ uncertainties of the inferred distributions.}
    \label{fig:simulation}
\end{figure}

@@

We next apply our hierarchical Bayesian framework to a sample of 161 exoplanetary systems with sky-projected stellar obliquity measurements, primarily consisting of Hot Jupiter systems, as summarized in \cite{Albrecht22} Table A1.
The inferred $\cos{\psi}$ distribution is shown in Figure~\ref{fig:psi_dist}. The $\cos{\psi}$ distribution is peaked at 1, with nearly flat behavior between $-0.75$ and $0.75$ and no significant clustering. The distribution suggests that there is a pile-up of planetary systems with stellar obliquities less than $40^\circ$ and an isotropic distribution for obliquities between $40^\circ$ and $140^\circ$.
The fraction of aligned systems dominates the distribution, with $w_0 = 0.67 \pm 0.09$. The corresponding hyperposteriors of the Beta distribution are $a_0 = 31.0\pm12.8$ and $b_0 = 0.40\pm0.12$. On the other hand, the fraction of misaligned systems is estimated to be $w_1 = 0.33 \pm 0.09$, with hyperposteriors of $a_1 = 1.56\pm0.93$ and $b_1 = 1.64\pm1.58$.

\begin{figure}[ht!]
    \script{psi_plot.py}
    \begin{centering}
        \includegraphics{figures/psi_dist.pdf}
        \caption{Inferred stellar obliquity distribution for all exoplanetary systems with sky-projected stellar obliquity measurements. This inference is based purely on the observed sky-projected obliquities and assumes an isotropic distribution for the stellar inclinations.}
        \label{fig:psi_dist}
    \end{centering}
\end{figure}

The discrepancy between the inferred stellar obliquity distribution of the full sample of exoplanetary systems and the previous analysis of the subsample with $i_\star$ measurements warrants further investigation. The earlier study identified a preponderance of perpendicular planets and disfavored an isotropic stellar obliquity distribution \citep{Albrecht21}. There are two potential explanations for this difference: (1) the subsample with $i_\star$ measurements is small and only includes about 20 misaligned systems, and (2) the requirement for $i_\star$ measurements could introduce selection biases in the sample.
